---
title: "R Notebook"
output: rmarkdown::github_document

---


# tidyYarns
tidyYarns is a function that binds the data from the types of yarns used in production

- Number of Millstyles
```{r}
length(unique(tidy_yarns$millstyle))
```

- Number of Yarns used in production 
```{r}
length(unique(tidy_yarns$yarn))
```

- types of yarn
```{r}
unique(tidy_yarns$type)
```

- number of purchased yarns
```{r}
unique_pur <- unique(tidy_yarns$yarn[which(tidy_yarns$type == 'Pur')])
num_pur <- length(unique_pur)
num_pur
```

- number of twisted yarns
```{r}
unique_twist <- unique(tidy_yarns$yarn[which(tidy_yarns$type == 'Twist')])
num_twist <- length(unique_twist)
num_twist
```

- number of dyed yarns
```{r}
unique_dye <- unique(tidy_yarns$yarn[which(tidy_yarns$type == 'Dye')])
num_dye <- length(unique_dye)
num_dye
```

- number of spun yarns 
```{r}
unique_spun <- unique(tidy_yarns$yarn[which(tidy_yarns$type == 'Spun')])
num_spun <- length(unique_spun)
num_spun
```

- Verify number of yarns
```{r}
num_pur + num_dye + num_twist + num_spun
```

# tidy_BOM details

- Number of MillStyles
```{r}
length(unique(tidy_Bom$millstyle))
```

- Number of Yarns

It appears that there is a 93 yarn difference between the bill of materials (tidy_Boms) and tidyYarns (630 unique yarns present in tidy_yarns)
```{r}
length(unique(tidy_Bom$yarn))
```

The yarns that differ between the two sets are:
```{r}
yarn_diff <- setdiff(tidy_Bom$yarn, tidy_yarns$yarn)
yarn_diff
```

- numbers of yarns in difference
```{r}
length(yarn_diff)
```

It appears there is a 173 yarn difference between the two datasets. 

Let's look into this...

```{r}
common_yarns <- unique(tidy_yarns$yarn[tidy_yarns$yarn %in% tidy_Bom$yarn])
length(common_yarns)
```

So based on the calculation above, only 550 yarns are present in both data sets.

```{r}
'%nin%' <- Negate('%in%')
tidy_yarn_diff <- tidy_yarns$yarn[tidy_yarns$yarn %nin% tidy_Bom$yarn]
unique(tidy_yarn_diff)
```
```{r}
length(unique(tidy_yarn_diff))
```


```{r}
'%nin%' <- Negate('%in%')
tidy_Bom_diff <- tidy_Bom$yarn[tidy_Bom$yarn %nin% tidy_yarns$yarn]
unique(tidy_Bom_diff)
```
```{r}
length(unique(tidy_Bom_diff))
```


- intesection between the tidy_yarns$yarn and tidy_Bom$yars
```{r}
tidy_bom_yarns_intersect <- intersect(tidy_yarns$yarn, tidy_Bom$yarn)
#tidy_bom_yarns_intersect
```

- number of common yarns between the two datasets
```{r}
length(tidy_bom_yarns_intersect)
```


- In regards to tidy_yarns$pounds, it is not specific as to what the amount listed per millstyle means


- Earliest Date Used Listed
```{r}
min(tidy_Bom$datelastused)
```

- Latest Date Used Listed
```{r}
max(tidy_Bom$datelastused)
```

# tidy_m365 Details

- number of millstyles present in m365

For reference, the number of unique millstyles in tidy_Bom is 957 and 535 for tidy_yarns
```{r}
length(unique(tidy_m365$millstyle))
```

- intersection between tidy_m365$millstyle and tidy_BOM$millstyle
```{r}
intersect(tidy_m365$millstyle, tidy_Bom$millstyle)
```



# tidyStops Details

- tidyStops only contains details for stoppage information for range and not millstyle

